---
title: "202C - HW1"
author: "Jiayu Wu"
date: "2018/4/29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### 1

$K_0$ is irreducible with paths connecting each two nodes, and it is aperiodic with period d=1.

$K_1$ is not irreducible since node 5 and 1 are not connected, and aperiodic because every state is aperiodic.

$K_2$ is irreducible, and not aperiodic with period d=2.

### 2
```{r}
K0 = matrix(c(.2, .8, .0, .0, .0, 
              .3, .0, .4, .0, .3,
              .1, .4, .0, .5, .0, 
              .0, .0, .4, .2, .4,
              .0, .0, .3, .6, .1),byrow = T, nrow = 5)
K1 = matrix(c(.3, .1, .3, .0, .3,
              .2, .4, .4, .0, .0,
              .0, .6, .3, .1, .0,
              .0, .0, .0, .5, .5,
              .0, .0, .0, .4, .6),byrow = T, nrow = 5)
K2 = matrix(c(.0, .0, .0, .3, .7,
              .0, .0, .0, .5, .5,
              .0, .0, .0, .8, .2,
              .0, .4, .6, .0, .0,
              .5, .0, .5, .0, .0),byrow = T, nrow = 5)
save(K0, file = "kernels.RData")
vals0 = eigen(K0)$values
vals1 = eigen(K1)$values
vals2 = eigen(K2)$values
vecs0 = eigen(t(K0))$vectors
vecs1 = eigen(t(K1))$vectors
vecs2 = eigen(t(K2))$vectors
d = cbind(cbind(rbind(vals0, vecs0), rbind(vals1, vecs1)), rbind(vals2, vecs2))
colnames(d) = c('K0_1','K0_2','K0_3','K0_4','K0_5','K1_1','K1_2','K1_3','K1_4','K1_5','K2_1','K2_2','K2_3','K2_4','K2_5')
rownames(d) = c('values','vectors','','','','')
knitr::kable(d, digits = 3, main="Eigen values and vectors")
```

### 3.

For $K_0$, there is one: $(0.096, 0.175, 0.245, 0.295, 0.189)$

For $K_1$, there is one: $(0, 0, 0, 0.444, 0.556)$

For $K_2$, there are two: $(0.092, 0.126, 0.282, 0.316, 0.184)$

### 4.
```{r}
load("kernels.RData")
power = diag(rep(1,5))
for (i in 1:200) {
      power = power %*% K0 
}
power
pi = power[1,]
save(K0, pi, file = "kernels.RData")
```
Yes, it does.

## Problem 2

### 1.
```{r}
load("kernels.RData")
v = c(0, 0, 1, 0, 0)
tv = c()
kl = c()
for (i in 1:200){
       v = v %*% K0
       tv[i] = 0.5*sum(abs(pi-v))
       kl[i] = sum(pi*log(pi/v))
}
plot(1:200, tv, type = "l", col="blue", xlab = "steps", ylab = "distance")
lines(1:200, kl, col = "green")
legend("topright", legend=c("TV_norm", "KL_div"), col = c("blue", "green"), lty = 1)
save(K0, pi, tv, file = "kernels.RData")
```

### 2.
```{r}
load("kernels.RData")
tvs = matrix(rep(0,25), nrow = 5)
for (i in 1:5) {
       for (j in i:5) {
              tvs[i,j] = 0.5*sum(abs(K0[i,]-K0[j,]))
       }
}
max(tvs)
# bound
boundsa = c()
power = diag(rep(1,5))
for (k in 1:100) {
       power = power %*% K0 
       tvs = matrix(rep(0,25), nrow = 5)
       for (i in 1:5) {
              for (j in i:5) {
                     tvs[i,j] = 0.5*sum(abs(power[i,]-power[j,]))
              }
       }
       boundsa[k] = max(tvs)
}
plot(1:100, boundsa, type = "l", xlab = "steps", ylab = "distance",
     main = "Contraction coeffcient bound C(K0)=1")
save(K0, pi, tv, boundsa,  file = "kernels.RData")
```


### 3.
```{r}
load("kernels.RData")
sq = sqrt((1-pi[3])/pi[3]/4)
boundsb = apply(t(1:100),2,function(x){sq*(0.61504869^x)})
plot(1:100, boundsb, type = "l", col = "blue", xlab = "steps", 
     ylab = "convergence", ylim=c(0,1), main = "Convergence rate plot")
lines(1:100, boundsa, col="green")
lines(1:100, tv[1:100], col="black")
legend("topright", legend=c("Bound A", "Bound B", "Convergence rate"), col = c("green", "blue", "black"), lty = 1)
# log-plot
plot(1:20, log(boundsb[1:20]), type = "l", col = "blue", xlab = "steps", 
     ylab = "convergence", main = "Convergence rate zoom-in log-plot")
lines(1:20, log(boundsa[1:20]), col="green")
lines(1:20, log(tv[1:20]), col="black")
legend("topright", legend=c("Bound A", "Bound B", "Convergence rate"), col = c("green", "blue", "black"), lty = 1)
```

## Problem 3
\begin{align*}
KL(\pi||v) -KL(\pi||\mu) &= KL(\pi(x)\mu(y)||\pi(y)\mu(x))=\sum_{x}\sum_{y}P(y,x)log\frac{\pi(x)\mu(y)}{\pi(y)v(x)}\\
&= \sum_{x}\sum_{y}P(y,x)log\frac{\mu(y)P(y,x)}{v(x)P(x,y)}  \\
&= \sum_{x}\sum_{y}P(y,x)log\frac{P(y,x)}{Q(y,x)} \\
&= E[KL(P(y, x)||Q(y, x))] \\
\\
\because P(y,x) &= \frac{\pi(x)P(x,y)}{\pi(y)} = \frac{\mu(y)\pi(x)}{v(x)\pi(y)}\\
KL(P(y,x)||P(y,x)) &= \sum_{x}\sum_{y}\frac{\pi(x)P(x,y)}{\pi(y)}log\frac{\pi(x)P(x,y)v(x)\pi(y)}{\pi(y)\mu(y)\pi(x)} \\
&= \sum_{x}\sum_{y}P(y,x)log\frac{P(x,y)v(x)}{\mu(y)} \\
&= \sum_{x}\sum_{y}P(y,x)logQ(y,x) = 0\\
E[KL(P(y, x)||Q(y, x))] &= \sum_{x}\sum_{y}P(y,x)log\frac{P(y,x)}{Q(y,x)}\\
&= \sum_{x}\sum_{y}P(y,x)logP(y,x)-\sum_{x}\sum_{y}P(y,x)logQ(y,x) \\
&= E_{P(y,x)}[logP(y,x)] \geq 0
\end{align*}


## Problem 4
$$P(\tau(0)<\infty)=\lim_{n\to\infty}\sum_{\tau(0)=1}^{n}P(\tau(0))=\lim_{n\to\infty}\sum_{\tau(0)=1}^{n}\alpha^{\tau(0)-1}(1-\alpha)=\lim_{n\to\infty}(1-\alpha)\frac{1-\alpha^{n-1}}{1-\alpha}=1$$

$$E[\tau_{ret}(0)]=\lim_{n\to\infty}\sum_{\tau(0)=1}^{n}\tau(0)P(\tau(0))=\lim_{n\to\infty}\sum_{\tau(0)=1}^{n}\tau(0)\alpha^{\tau(0)-1}(1-\alpha)=(1-\alpha)\lim_{n\to\infty}(\frac{1-\alpha^n}{(1-\alpha)^2}+\frac{n\alpha^n}{1-\alpha})=(1-\alpha)\frac{1}{(1-\alpha)^2} = \frac{1}{1-\alpha}$$
